---
title: "Data Taming R reminder sheet"
author: "Bill S. Preston Esq."
date: "`r Sys.Date()`"
output:
  pdf_document:
    number_sections: true
---

# Setup

## Initialising `knitr`

* This first code chunk will execute, but not display in the output because of the option `include=FALSE` in the definition of the code chunk.

`knitr` general options:

* `message=FALSE` when setting the `knitr` general options to suppress messages when loading the packages later in the file.
* `results=FALSE` to suppress output. (Change this to `results=TRUE` to see the output.)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, results=FALSE, warning=FALSE)
```

## Loading packages

```{r tidyverse}
library(tidyverse)
library(inspectdf)
library(lubridate)
library(caret)
library(moments)
library(tidymodels)
library(ISLR)
library(car)
```

# Loading data

* Use `<-` to assign new datasets and variables

## Data already in the library

```{r}
data("mpg")
mpg1<-mpg
```


## Read in a `csv` file

```{r}
pop1<- read_csv("population.csv")
```

# Displaying data

## Displaying data frames and tibbles

```{r}
mpg1 # Prints the first 10 lines
print(n=10, mpg) # Also prints the first 10 lines
head(mpg1, 9) # Also prints the first 9 lines
tail(mpg1, 8) # Prints the final 8 lines
View(mpg1) # This command will display data in a new window
```

## To show the number of rows and columns of the data

* `dim()` gives a list of the form `[\# rows, \# columns]`

```{r}
dim(mpg)
```

* To just find the number of rows use `nrow()`

```{r}
nrow(mpg)
```

* To just find the number of columns use `ncol()`

```{r}
ncol(mpg)
```


## To access a particular column

```{r}
mpg1$cyl
```

# Extracting data

## Extracting columns
```{r}
select(mpg, cyl:fl)
```

## Extracting specific row numbers
```{r}
mpg[5,]
```

## Extracting specific values

### Just keeping the value

```{r}
mpg[5,]$trans
class(mpg[5,]$trans)
```

### Putting the value into its own new tibble

```{r}
mpg[5,"trans"]
class(mpg[5,"trans"])
```

## Extracting rows that match a TRUE/FALSE condition
```{r}
filter(mpg, displ==3.1)
filter(mpg, between(mpg$displ,2.8,3.1))
```

# Missing data

## To check if there are any missing values in any columns

```{r}
inspect_na(starwars)
```

## Finding missing values in a specific column

* This returns a boolean list of TRUE/FALSE indicating the rows with missing data.

* This can be combined with the filter() command

```{r}
is.na(starwars$species)
```


# Manipulating data

## Sorting a column

```{r}
arrange(mpg, displ)
```


## `mutate()` To add, change or remove columns

* Add column to right of dataset
  - `mutate(dataset, new_column_name = value)`

```{r}
mutate(mpg1, IDnum=c(1:234))
mutate(mpg1, cty_hwy_avg=(cty -hwy)/2)
```

* Delete a column
  - `mutate(dataset, existing_column_name = NULL)`

```{r}
mutate(mpg1, model=NULL)
```

* Change a column
  - `mutate(dataset, existing_column_name = value)`

```{r}
mutate(mpg, displ=displ*10)
```

## `rename()` to rename a column

* Use the syntax `rename(dataset, new_column_name=old_column_name)`

```{r}
rename(mpg, displacement=displ)
```


## `relocate()` to move a column

* Move a column "before" (to the left) of another column

```{r}
relocate(mpg, "cyl", .before = model)
```

* Move a column "after" (to the right) of another column

```{r}
relocate(mpg, "cyl", .after = cty)
```

## Grouping rows

* Group a set of rows together based on the values in one of the columns. Eg. this will group all the cars together by their number of cylinders.

* This can be used with the `summarise()` command to computer statistics for each group.

```{r}
group_by(mpg,cyl)
```



## Change data types

* Convert to factor (nominal categorical variable)

```{r}
mpg1$cyl<-as.factor(mpg1$cyl)
```

* Convert to ordered factor (ordinal categorical variable)

```{r}
mpg1$cyl<-as.ordered(mpg1$cyl)
```

* Convert to numerical variable

```{r}
mpg1$cyl<-as.numeric(mpg1$cyl)
```

* Convert to integer variable

```{r}
mpg1$cyl<-as.integer(mpg1$cyl)
```

* Convert to character string

```{r}
mpg1$cyl<-as.character(mpg1$cyl)
```

* Convert to logical/Boolean variable

  - First need a column of TRUE/FALSE or 1/0

```{r}
mpg2<-mutate(mpg1, tf="TRUE")
mpg2$tf<-as.logical(mpg2$tf)
```

* Convert to date object

  - Using `lubridate` package commands `ymd` or `dmy`

```{r}
mpg2<-mutate(mpg1,date="2025-10-05")
mpg2$date<-ymd(mpg2$date)
mpg2<-mutate(mpg1,date="05-10-2025")
mpg2$date<-dmy(mpg2$date)
```

## Rename entries in column

* Using `fct_recode(mpg$drv, "new1"="old1", "new2"="old2", "new3"="old3")`
* Only works when column is a factor or character string
* Also converts column to `factor` type

```{r}
fct_recode(mpg$drv, "front"="f", "4x4"="4", "rear"="r")
```

# A sequence of numbers

```{r}
1:50
```


# Random sampling

## Random sampling from a tibble

* Use `sample_n()`

* Uniformly at random choose 20 rows with replacement

```{r}
sample_n(mpg, 20, replace=TRUE)
```

* Uniformly at random choose 20 rows without replacement

```{r}
sample_n(mpg, 20, replace=FALSE)
```

## Generating a list of random numbers

* Use `sample()`

* Generate 7 random integers from 1 to 50 with replacement

```{r}
sample(1:50, 7, replace = TRUE)
```

* Generate 7 random integers from 1 to 50 without replacement

```{r}
sample(1:50, 7, replace = FALSE)
```

## Setting the seed for a random number generation

```{r}
set.seed(1234)
```


# Character string manipulation

## Special characters

When using regular expressions you need the following commands for special characters:

* \$: use `\\$`

* (: use `\\(`


## Joining (concatenating) strings

* These commands `paste0` and `str_c` seem to do the same thing

```{r}
middle<-"middle bit;"
paste0("first bit;",middle, " last bit")
str_c("first bit;", middle, " last bit")
```


## Extracting numbers from strings

* To extract numbers from strings use regular expressions. Eg. `(\\d+)`

```{r}
df  <- tibble(
  treatment = c("A", "B", "C"), 
  response = c(12, 11, 10),
  some_text1 = c("abc 7", "abc 2", "abc 5"),
  some_text2 = c("abc 7 xyz 9", "abc 1 xyz 21", "abc 0 xyz 2"),
  some_text3 = c("abc 7 xyz 9", "abc -2 xyz 21", "abc 0.5 xyz 2")
)
str_match(df$some_text1, "abc (\\d+)")
```

* Also works for extracting multiple values from a string

```{r}
str_match(df$some_text2, "abc (\\d+) xyz (\\d+)")
```

* Only works for positive integers

```{r}
str_match(df$some_text3, "abc (\\d+) xyz (\\d+)")
```

## Extracting alphabetic characters from strings

* Works for both upper and lower case letters

```{r}
str_match(df$some_text2, "x([:alpha:]+)")
```


## Replacing parts of strings

```{r}
mpg[6,]$trans
str_replace(mpg[6,]$trans,"a","X")
str_replace(mpg[6,]$trans,"\\(","X")
```

### To replace all parts of the string matching the pattern

```{r}
mpg[6,]$trans
str_replace_all(mpg[6,]$trans,"a","X")
```

# Precision

* Rounding off to $n$ decimal places
  - Note that $n$ can be zero or negative. (Experiment with it to see what it does.)

```{r}
round(15.32257,3)
round(15.32257,-1)
```

* Rounding off to $n$ significant figures

```{r}
signif(15.32257,3)
```


# Statistics

* Mean

```{r}
mean(mpg$hwy)
```

* Sample standard deviation. (Note that the sample deviation uses $N-1$ in the denominator of the calculation.)

```{r}
sd(mpg$hwy)
```

* Skewness

```{r}
skewness(mpg$hwy)
```

## Statistics on grouped data

* We can use the statistics commands via the `summarise()` command, which is especially useful for working with grouped data

```{r}
summarise(mpg, mean_hwy = mean(hwy))
summarise(group_by(mpg,cyl), mean_hwy = mean(hwy))
```


## Summary statistics

* Calculate summary statistics for all numerical variables we can use `inspect_num()`

```{r}
inspect_num(mpg)
```



## Building formulae

* To write a formula we put the response variable on the left of $\sim$ and the predictors on the right. Eg. `y ~ x + z`.

* We can include interaction between our predictors with the colon ":". Eg. ` y ~ x + z + x:z`

* The easy way to write a formula with all individual terms and second-order interactions as predictors is `y ~ .^2`.
  - For example, if we have a set of predictors `x1, x2, x3` then `y ~ .^2` is equivalent to `y ~ x1 + x2 + x3 + x1:x2 + x1:x3 + x2:x3`

## Linear models

```{r}
df<- tibble(
  x=c(1:20),
  y=x+rnorm(20,0,2)
)
df_lm<-lm(y~x,df)
summary(df_lm)
```

* Extracting the model coefficients

```{r}
as.numeric(df_lm$coefficients)
```


### Graphs for checking assumptions of linear models

```{r}
plot(df_lm, which=1)
plot(df_lm, which=3)
plot(df_lm, which=2)
```

### Predicting with linear models

**(Note that the explanatory values used for the prediction must be stored in a tibble/dataframe.)**

* Prediction with prediction interval of level 85\%

```{r}
pred_values<-tibble(
  x=c(2.5, 7.2)
)
predict(df_lm, pred_values, interval="prediction", level = 0.85)
```

* Prediction with confidence interval of level 99\%

```{r}
predict(df_lm, pred_values, interval="confidence", level = 0.99)
```

## Logistic models

* We first make a binary categorical variable in a data set

```{r}
car_seats <- as_tibble(Carseats)
car_seats
car_seats <- car_seats %>% 
  mutate("sales_high"=ifelse(Sales>8,"high","low"), .after = Sales)
car_seats$sales_high <- factor(car_seats$sales_high)
car_seats_1 <- car_seats %>% 
  mutate(Sales=NULL)
```

### Building the model

```{r}
classification_lr <- logistic_reg() %>%
  set_engine("glm")
lrfit <- classification_lr %>%
  fit(sales_high ~ Price, data = car_seats_1)
```

### Predicting with the logistic model

```{r}
predict(lrfit, new_data=car_seats_1)
predict(lrfit, new_data=car_seats_1, type="prob")
```

## Extracting the model data for general linear models

* Summary of the model

```{r}
summary(lrfit$fit)
```

* Just the coefficients

```{r}
as.numeric(lrfit$fit$coefficients)
```

## Analysis of variance

```{r}
Anova(lrfit$fit)
```


# Data transforms

## Box-Cox transform

### Finding $\lambda$ value

```{r}
df_bc<-BoxCoxTrans(y=df$y, x=df$x)
df_bc$lambda
```

### Transforming the data

```{r}
predict(df_bc,df$y)
```



# Manipulating time

* Calculate the duration between two date objects in days

```{r}
dmy("05-11-2028")- dmy("05-10-2025")
```

* Calculate the duration between two date objects in seconds

```{r}
as.duration(dmy("05-11-2028")- dmy("05-10-2025"))
```

# Cleaning data

## Finding duplicates

```{r}
df<- tibble(
  x=c(1:7,5,2),
  y=x^2
)
duplicated(df)
```
(To use this to extract the duplicated values, see Section "Extracting rows that match a TRUE/FALSE condition".)


# Tidying data

* Convert to "long" form ("melting" your data)
  - `gather(dataset, key= "first_new_column", value = "second_new_column", start:end)`

```{r}
data("table4a")
table4a
TB_cases<- gather(table4a, key = "year", value = "cases", `1999`:`2000`)
```

* Convert back to "wide" form, where
  - `key` is the variable that you want to put as new column headings
  - `value` is the the variable that you want to put in these new columns
  - `spread(dataset, key= "key_column", value = "data_column")`

```{r}
spread(TB_cases, key = "year", value = "cases")
```
# Plotting

* Scatter plot

```{r}
ggplot(mpg, aes(x=displ, y=cty))+
  geom_point()
```

* Line plot

```{r}
ggplot(mpg, aes(x=displ, y=cty))+
  geom_line()
```

* Box plot

```{r}
ggplot(mpg, aes(class, cty))+
  geom_boxplot()
```

* Bar graph

```{r}
ggplot(mpg, aes(x = manufacturer)) + geom_bar()
```

This is similar to a histogram (but we should use bar charts for categorical variables, and histograms for quantitative variables)

```{r}
ggplot(mpg, aes(manufacturer))+
  geom_histogram(stat="count")
```

## Rotating x-axis label

```{r}
ggplot(mpg, aes(x = manufacturer))+
  geom_bar()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

## Fitting lines and curves to the data

* A straight trend line. Need the method "lm" (as in "linear model")

```{r}
ggplot(mpg, aes(x=displ, y=cty))+
  geom_point()+
  geom_smooth(method="lm")
```

* A possibly curved smoothed average line

```{r}
ggplot(mpg, aes(x=displ, y=cty))+
  geom_point()+
  geom_smooth()
```

# R Markdown

## Writing equations

* R Markdown uses Latex conventions for equations (so if you know Latex, you can just type it directly into R Markdown)

* To write in-line maths (expressions in the middle of the text) use a single dollar \$ followed by your maths then a second single dollar \$. Eg. this equation $y=mx+c$ is an in-line maths expression.
  - Note that `\(...\)` is equivalent to `$...$` and you can use either.
  - Note that this means you can't just type a dollar symbol in R Markdown. To type a dollar symbol use `\$`. Eg. a \$10 note is blue.

* To write "displayed" maths (expressions separated from the text) use a double dollar \$\$ followed by your maths, then another double dollar \$\$. Eg. this equation $$y=x^3+4x+1$$ is a "displayed" expression.
  - Note that `\[...\]` is equivalent to `$$...$$` and you can use either. 

## Writing long equations

* We can use the `align` environment.
* You end a row in the equation with \\\\
* Use an ampersand (&) to align the rows 
* This also puts an equation number on every row of our equation

\begin{align}
y&= \alpha x^3 +4 \beta x + 3\gamma\\
&+ 321 - \omega\\
&= 789
\end{align}

* If you don't want the equation number, use the `align*` environment

\begin{align*}
y&= \alpha x^3 +4 \beta x + 3\gamma\\
&+ 321 - \omega\\
&= 789
\end{align*}

## Writing maths symbols (including Greek letters)

* The easiest way to write maths symbols, including Greek letters, is to write them as in-line maths. Eg. here is an $\alpha$, here is a $\beta$ and here is a $\gamma$.
  - To type a Greek letter, type a backslash followed by the name of the letter, eg. `\alpha`, `\beta`, `\gamma`.

* To put a "hat" on a maths symbol write `\hat{<symbol>}`. Eg. here is a regular $\alpha$ and here is alpha-hat $\hat{\alpha}$.

* To write subscripts use an underscore{} after the symbol `<symbol>_{subscript}`. Eg. here is a regular $\alpha$ and here is $\alpha_{1,2}$
  - If you've only got a single character in the subscript, then you don't need the curly brackets `{}`, but it's good practice to use them anyway. Eg. here is a regular $\alpha$ and here is $\alpha_1$ (no curly brackets) and here is $\alpha_{1}$ (with curly brackets).

### Some other symbols

* Approximation: `\approx`, ie. $\approx$



## Including R output in your text

* To include some numerical output from R directly into your knitted text use `` `r knitr::inline_expr("<variable name>")` ``

```{r}
v_1 <- 10.5324
```

Then we can print the value of the variable `r v_1`.

* To include the numerical output in an an equation, just do the same thing inside \$, or \$\$ or the `align` environments

Our variable is $v_1=`r v_1`$, or we could write
\begin{align*}
y &= v_1 \alpha \\
&= `r v_1` \alpha 
\end{align*}
