---
title: "Data Taming Assignment 3"
author: "Dongju Ma"
date: "2024-08-10"
output:
  pdf_document:
    default
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

```{r}
library(tidyverse)
library(tidymodels)
library(car)
library(modelr)
```

# Q1. Loading the data

```{r}
# Your student number goes here
ysn = 1942340
# Calculate your student number modulo 3
filenum <- (ysn + 2) %% 3
filenum
filename <- paste0("./data/merry_",filenum,".csv")
filename
# Read the csv file
merry <- read_csv(filename)
merry
# Output the dimensions
dim(merry)
```

# Q2. Identifying data types

-   RHBMM: Categorical Nominal\
    This column contains only two types of values: 1 and 0, representing whether the archer is married or not. Since these values are not ordered, this data should be classified as Categorical Nominal
-   Accuracy: Quantitative Continuous\
    This column represents the proportion of successful hits out of all attempts, with values ranging continuously between 0 and 1. So it should be classified as Quantitative Continuous.
-   AGE: Categorical Ordinal\
    This column categorizes archers into three age groups, ordered sequentially from youth to middle age, and finally to senior. So it should be classified as Categorical Ordinal.
-   DRESS: Categorical Nominal\
    This column shows the colors of each archer's outfit. There are three colors and the colors are not ordered, so it should be classified as Categorical Nominal.
-   Home: Categorical Nominal\
    This column only contains two kinds of values which are forest and city. Between these values there are no special orders. So it should be classified as Categorical Nominal.
-   Jail: Categorical Nominal\
    This column contains yes and no these two types of values which represents the archer was in jail before or not. These values are not ordered either so it still should be Categorical Nominal.

# Q3. Taming Data

```{r}
# Change the column names
merry <- rename(merry,  rhbmm = RHBMM, acc = Accuracyyy, age = AGE, dress = DRESS,
                home = Home, jail = Jail)
# Transform into factors
merry$rhbmm <- as.factor(merry$rhbmm)
merry$age <- as.factor(merry$age)
merry$dress <- as.factor(merry$dress)
merry$home <- as.factor(merry$home)
merry$jail <- ifelse(merry$jail == 'yes', TRUE, FALSE)
merry$jail <- as.logical(merry$jail)
# Show the final version and the dimensions
merry
dim(merry)
```

# Q4. Set the training set and testing set

```{r}
# Set seed
set.seed(1942340)
# Split the data set
merry_split <- initial_split(merry, prop = 2/3)
merry_train <- training(merry_split)
merry_test <- testing(merry_split)
# Show the results
merry_split
merry_train
dim(merry_train)
merry_test
dim(merry_test)
```

# Q5. Build a logistic regression model

```{r}
classification_lr <- logistic_reg() %>%
  set_engine('glm')
# Fit a logistic regression model with each variable individually
lrfit <- classification_lr %>%
  fit(rhbmm ~ acc+age+dress+home+jail, data = merry_train)
summary(lrfit$fit)
```

# Q6. Fit a matrix model

```{r}
model_matrix(merry_train, ~dress+age)
```

## Q6.(a) New Variables

For dress variable, there are two new variables introduced in the model.One is dressgreen and the other is dressred. For age variable there are also two new variables introduced, one is agesenior and the other one is ageyouth.

## Q6.(b) Reference levels

The reference level for dress is black, and the reference level for age is middle.

# Q7. Explanations for the logistic regression models

I think there should be 10 separate lines in this model, since we use logistic regression to each variable individually. We would have lines represent the relationships between the only continuous predictor accuracy and the categorical predictors, which are age, dress, home and jail. Age has three categories and so does dress. The last two variables each has two categories. So after all the each category has one separated line with the accuracy. So there should be 10.

# Q8. Model with Interaction terms

```{r}
# Fit a model with all second-order interaction terms
lrfit0 <- classification_lr %>%
  fit(rhbmm ~ acc+age+dress+home+jail+
        acc:age+acc:dress+acc:home+acc:jail+
        age:dress+age:home+age:jail+
        dress:home+dress:jail+
        home:jail,
      data = merry_train
      )
# Use Anova to evaluate the p-values
Anova(lrfit0$fit)
```

As the output above, acc:home and dress:home meet the 99% significance level.

# Q9. Backwards stepwise regression

## Q9.(a) New model without Jail

```{r}
# First fit a new model without jail interactions
lrfit1 <- classification_lr %>%
  fit(rhbmm ~ acc+age+dress+home+jail+
        acc:age+acc:dress+acc:home+
        age:dress+age:home+
        dress:home,
      data = merry_train
  )
Anova(lrfit1$fit)
```

As the output above we can learn that the interaction terms acc:dress, age:dress and age:home cannot meet the 95% significance level, so we just remove them and fit a new model.

## Q9.(b) Clean the interaction terms

```{r}
# Fit a model with the interaction terms removed
lrfit2 <- classification_lr %>%
  fit(rhbmm ~ acc+age+dress+home+jail+
        acc:age+acc:home+
        dress:home,
      data = merry_train
  )
Anova(lrfit2$fit)
```

As the output above, the age variable doesn't meet 95% significance level, so we should remove it and its interactions because of the “principle of marginality” next.

## Q9.(c) Clean the individual terms

```{r}
# Fit a model with age removed
merry_lr_fit <- classification_lr %>%
  fit(rhbmm ~ acc+dress+home+jail+acc:home+dress:home,
      data = merry_train
  )
Anova(merry_lr_fit$fit)
```

Now we have a model with only individual variables and interaction terms with 95% significance level.

# Q10. Analysis the interactions

## Q10.(a) Interaction terms

The interaction between accuracy and home and the interaction between dress and home are significant.

## Q10.(b) Hypotheses

I think the archer's residence determines their accuracy of archery and their dress color. The forest archers are more likely to have a higher accuracy and dress themselves in protect colors like green. Also these interactions may affect whether they are married or not, the forest archers with higher accuracy and with green outfits dressed may have less chance to get married.

# Q11. The equation for log-odds

The equation of $\hat{r}_i$ should be, $$
\hat{r}_i = \hat{\beta_0} + \hat{\beta_1}x_1 + \hat{\beta_2}x_2 + \hat{\beta_3}x_3 
+ \hat{\beta_4}x_4 +\hat{\beta_5}x_5 + \hat{\beta_6}x_1x_4 + \hat{\beta_7}x_2x_4 +\hat{\beta_8}x_3x_4
$$ As the output above and for this equation,\
$\hat{\beta_0}$ means the intercept with the reference under the circumstances of dressing in black, living in the city and having not been in jail before.\
$x_1$ means the log-odds of accuracy predictors and $\hat{beta_1}$ is its slope.\
$x_2$ means the log-odds under dressing green and $\hat{beta_2}$ is its slope.\
$x_3$ means the log-odds of dressing in red and $\hat{beta_3}$ is its slope.\
$x_4$ means the log-odds of living in forest and $\hat{beta_4}$ is its slope.\
$x_5$ means the log-odds of being in jail before and $\hat{beta_5}$ is its slope.\
$\hat{\beta_6}x_1x_4$ is the interaction term of accuracy and living in the forest with its coefficient.\
$\hat{\beta_7}x_2x_4$ is the interaction term of dressing green and living in the forest with its coefficient.\
$\hat{\beta_8}x_3x_4$ is the interaction term of dressing red and living in the forest with its coefficient.

# Q12. Find the geometric meaning of the model

## Q12.(a) Separate lines

We have 8 lines of this model. Because we have 8 terms in the equations with predictors, each term means there should be a line existing.

## Q12.(b) Are they paralleled?

They are not paralleled as they have different coefficients from each other which means the slopes of every line is different.

# Q13. Fit the equation with model coefficents

```{r}
# Show the summary of the latest regression model
summary(merry_lr_fit$fit)
```

As the output above and what we described in question 11, we could put the coefficients into their positions where they should be.\
So the equation could also be written like, $$
\hat{r}_i = -1.302 + 1.29x_1 + 0.897x_2 - 0.177x_3  + 0.624x_4 +0.175x_5 + 2.04x_1x_4 + 2.05x_2x_4 - 1.079x_3x_4
$$

# Q14. Estimations with the model

## Q14.(a) Esitimation 1

With an archer who is middle aged, lives in the forest, wears red clothing and has been to jail we should calculate with intercept plus $- 0.177x_3$, $0.624x_4$, $0.175x_5$ and $- 1.08x_3x_4$, with $x_3, x_4, x_5$ all being 1.\
Which should be $$
\begin{aligned}
&-1.302 - 0.177x_3 + 0.624x_4 + 0.175x_5 -1.079x_3x_4\\
&=
-1.302 - 0.177 + 0.624 + 0.175 - 1.079 \\
&= 
-1.759
\end{aligned}
$$

## Q14.(b) Esitimation 2

With an archer who is quite old, lives in the city, usually wears black clothing and has never been to jail, we should calculate with just intercept. So it should be -1.302

# Q15. Testing the model

```{r}
# Create a new prediction tibble with true class and predicted class
merry_test_preds <- bind_cols(
  predict(merry_lr_fit, new_data = merry_test),
  truth = merry_test$rhbmm,
  predict(merry_lr_fit, new_data = merry_test, type = 'prob'),
)
merry_test_preds
# Show the dimensions of the tibble
dim(merry_test_preds)
```

# Q16. Evaluating the model

## Q16.(a) Confusion matrix and accuracy

Find the confusion matrix and the accuracy of the model.

```{r}
# Build the confusion matrix
merry_test_cm <- merry_test_preds %>%
  conf_mat(
    .pred_class,
    truth = truth
  )
merry_test_cm
```

The confusion matrix is above.

```{r}
# Calculate the accuracy
merry_test_preds %>% accuracy(
  .pred_class,
  truth = truth
)
```

The accuracy of this model is 0.662.

## Q16.(b) Sensitivity and specificity

We would calculate the sensitivity first. It should be the true positives proportion of the sum of true positives and false negatives.\
Which should be $\frac{3548}{2052+3548} \approx 0.634$\
Then the specificity, which should be the true negatives proportion of the sum of true negatives and false positives. Which should be $\frac{3076}{3076+1324} \approx  0.699$

## Q16.(c) ROC plot

```{r}
# Plot the roc curve with probability predictions
merry_test_preds %>%
  roc_curve(
    .pred_0,
    truth = truth
  ) %>%
  autoplot()+
  geom_vline(xintercept = 0.5, linetype = 'dashed') +
  geom_hline(yintercept = 0.5, linetype = 'dashed')
```

## Q16.(d) AUC-ROC

```{r}
merry_test_preds %>%
  roc_auc(
    .pred_0,
    truth = truth,
    event_level = 'first'
  )
```

The AUC of the ROC curve is 0.721.

# Q17. Conclusions

AS we know the young lad has an accuracy of 112/116, he haven't been in jail before, he's wearing a green outfit and lives in the forest. According to the model before we could make a prediction like this,

```{r}
# create a tibble to store the target's information
target <- tibble(
  acc=112/116,age='youth',dress='green',home='forest',jail=FALSE
)
# Predict with the model
target_preds <- bind_cols(
  target,
  predict(merry_lr_fit,new_data = target),
  predict(merry_lr_fit,new_data = target, type = 'prob')
  )
# Show the result
target_preds
```

So according to the output the new archer is very likely to be a member of Robin Hood’s Merry Men, with about 99.60%.
